Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-22 paper/abstract.tex
Add todos and fixed spelling
fiddling with contracts discussion

Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they either gracefully recover or safely shut down as requirements dictate. The DARPA Cyber Assured Systems Engineering program is developing tools for design, analysis, and verification that enable systems engineers to design-in cyber-resiliency in a Model-Based Systems Engineering environment. This paper describes automated model transformations that introduce high-assurance cyber-resiliency components into a system, in particular filters and monitors that prevent malicious input and detect supply chain attacks, respectively. A formal specification defines each high-assurance component, and is used to verify that the component addresses system level cyber requirements. Implementations for these high-assurance components are directly synthesized from their specifications, and are automatically proven to preserve the exact meaning of the specifications all the way down to the binary code level. The model transformations are integrated into the Open Source AADL Tool Environment (OSATE). The paper further reports on a case study applying security-enhancing model transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning. In the case study, the model transformations add filters to guard against malformed input, as well as monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

>>>>

Cyber-physical systems, such as avionics, must be tolerant to
cyber-attacks in the same way they are tolerant to random faults: they
either gracefully recover or safely shut down as requirements dictate.
The DARPA Cyber Assured Systems Engineering program is developing
tools for design, analysis, and verification that enable systems
engineers to design-in cyber-resiliency in a Model-Based Systems
Engineering environment.  This paper describes automated model
transformations that introduce high-assurance cyber-resiliency
components into a system, in particular filters and monitors that
prevent malicious input and detect supply chain attacks, respectively.
A formal specification defines each high-assurance component, and is
used to verify that the component addresses system level cyber
requirements.  Implementations for these high-assurance components are
directly synthesized from their specifications, and are automatically
proven to preserve the exact meaning of the specifications all the way
down to the binary code level.  The model transformations are
integrated into the Open Source AADL Tool Environment (OSATE).  The
paper further reports on a case study applying cyber-resiliency model
transformations to a UAV system that uses the Air Force Research
Laboratory's OpenUxAS services for route planning.  In the case study,
the model transformations add filters, to guard against malformed
input, as well as monitors, to guard against ground station spoofing
and malicious flight plans from OpenUxAS.

### 2022-03-22 paper/intro.tex
Add todos and fixed spelling
fiddling with contracts discussion

BriefCASE incorporates model-level cyber analysis tools (presently
GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
examine AADL models for potential vulnerabilities and suggest
cyber-security requirements to mitigate them.  A library of
architectural transforms guides the system engineer through automated
model transformations that modify the architecture to address these
requirements, possibly inserting new high-assurance components into
the system.  Implementations for the new components are synthesized
from formal specifications using
SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
(Semantic Properties for Language and Automata Theory).  Formal
verification that the transformed system model meets its requirements
is accomplished via AGREE~\cite{agree2013} (Assume Guarantee Reasoning
Environment).
%The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
AGREE is a {\em compositional assume-guarantee} style model checker
for AADL models that attempts to prove properties about one layer of
an architecture using properties allocated to its subcomponents.
Cyber-resilient code implementing the verified model is then
automatically generated using the High Assurance Modeling and Rapid
Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
desired, this code can be targeted to the formally verified seL4
secure microkernel~\cite{sel4-2009}.

@@

Two security-enhancing transformations are discussed: (1) the
insertion of a filter to prevent malformed data from a malicious actor
from being propagated to downstream components, and (2) the insertion
of a monitor to detect (and alert) unexpected behaviors arising from
untrusted components. These transformations not only change the
architecture of the model by adding in new components; they also
generate a formal specification for the behavior of the inserted
high-assurance components in the AGREE language. Those specifications
are sufficient for model checking to prove that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

>>>>

BriefCASE incorporates model-level cyber analysis tools (presently
GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) which can
examine AADL models for potential vulnerabilities and suggest
cyber-security requirements to mitigate them.  A library of
architectural transforms guides the system engineer through automated
model transformations that modify the architecture to address these
requirements, possibly inserting new high-assurance components into
the system.  Implementations for the new components are synthesized
from formal specifications using
SPLAT~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}
(Semantic Properties for Language and Automata Theory).

Formal
verification that the transformed system model meets its requirements
is accomplished via AGREE~\cite{agree2013} (Assume Guarantee Reasoning
Environment).
%The AGREE analysis \emph{assumes} properties on the inputs of a given component of the system, and attempts to formally prove the conjectured \emph{guarantees} of the output.
AGREE is a {\em compositional assume-guarantee} style model checker
for AADL models that attempts to prove properties about one layer of
an architecture using properties allocated to its subcomponents.
Cyber-resilient code implementing the verified model is then
automatically generated using the High Assurance Modeling and Rapid
Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If
desired, this code can be targeted to the formally verified seL4
secure microkernel~\cite{sel4-2009}.

@@

Two cyber-resiliency transformations are discussed: (1) the insertion
of a filter to prevent malformed data from a malicious actor from
being propagated to downstream components, and (2) the insertion of a
monitor to detect (and alert) unexpected behaviors arising from
untrusted components. These transformations not only change the
architecture of the model by adding in new components; they also
generate a formal specification for the behavior of the inserted
high-assurance components in the AGREE language. Those specifications
are sufficient for model checking to prove that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

### 2022-03-22 paper/verification-conditions.tex
fiddling with contracts discussion

A \emph{system} $\system = (\inputs, \outputs, \assumes,
\guarantees)$, where $\inputs$ is the input set, $\outputs$ is the
output set, $\assumes$ is the set of assumptions, and $\guarantees$ is
the set of guarantees.  A system is implemented by a set of
\emph{components}, $\components$, together with with their
inter-connections.  A component $\component$ is defined by its own
tuple in a manner similar to the system as $(\inputs_\component,
\outputs_\component, \assumes_\component, \guarantees_\component)$.
From the components and connections, $\mathbb{I}_\component$ is
defined to be the set of components providing input to some component
$\component$ in the system, and $\mathbb{O}$ is defined to be the set
of components that provide the output for the system.  A system is
correct if and only if for all $\pi$ and for all $i \ge 0$ the
following conditions hold:
\[
\begin{array}{lll}
        & \forall \component \in \components &
            \globally(\historically(\assumes \wedge
            \bigwedge_{\component^\prime \in \mathbb{I}_\component} P_{\component^\prime})
            \implies \assumes_\component) \\
 \wedge &   &
            \globally(\historically(\assumes \wedge
            \bigwedge_{\component^\prime \in \mathbb{O}} \guarantees_{\component^\prime})
            \implies \guarantees)
\end{array}
\]
The first condition verifies the input assumptions on each component
under the system assumptions and upstream component guarantees.  It
checks if the component guarantees and system assumptions are strong
enough to imply input assumptions on all immediate downstream
components.  The second condition checks the output guarantees of the
system under the system assumptions and component guarantees that
provide the output.  It checks if the guarantees on components
providing primary outputs are strong enough to imply the system
guarantees.

>>>>

A \emph{system} $\system = (\inputs, \outputs, \assumes,
\guarantees)$, where $\inputs$ is the input set, $\outputs$ is the
output set, $\assumes$ is the set of assumptions, and $\guarantees$ is
the set of guarantees.  A system is composed of a set of
\emph{components}, $\components$, together with their
inter-connections.  A component $\component$ is, hierarchically, also
a system, and may be designated by its own tuple
$(\inputs_\component, \outputs_\component, \assumes_\component,
\guarantees_\component)$.  From the components and connections,
$\mathbb{I}_\component$ is defined to be the set of components
providing input to some component $\component$ in the system, and
$\mathbb{O}$ is defined to be the set of components that provide the
output for the system.  A system $S$ is \emph{correct} if and only if for all
components $c \in C$:
\begin{equation}
            \globally(\historically(\assumes \wedge
            \bigwedge_{\component^\prime \in \mathbb{I}_\component} P_{\component^\prime})
            \implies \assumes_\component)
\end{equation}
\begin{equation}
            \globally(\historically(\assumes \wedge
            \bigwedge_{\component^\prime \in \mathbb{O}} \guarantees_{\component^\prime})
            \implies \guarantees)
\end{equation}
Condition (1) verifies the input assumptions on each component under
the system assumptions and upstream component guarantees.  It checks
if the component guarantees and system assumptions are strong enough
to imply input assumptions on all immediate downstream components.
Condition (2) checks the output guarantees of the system under the
system assumptions and component guarantees that provide the output.
It checks if the guarantees on components providing primary outputs
are strong enough to imply the system guarantees.
